---
title: "MoneyWicket"
author: "Joseph Ellis, Abhiram Reddy"
subtitle: 
output:
  html_document:
    df_print: paged
  html_notebook: default
---

```{r setup, include=FALSE}

library(knitr)

set.seed(2077)
options(width=70)
options(scipen=99)

opts_chunk$set(tidy.opts=list(width.wrap=50),tidy=TRUE, size = "vsmall")  
opts_chunk$set(message = FALSE,                                          
               warning = FALSE,
               cache = FALSE,                               
               autodep = TRUE,
               cache.comments = FALSE,
               collapse = TRUE,
               fig.width = 5.5,  
               fig.height = 4.5,
               fig.align='center')

```

```{r setup-2}

sessionInfo()
getwd()

```

```{r setup-3}

library('tidyverse')
library('readr')
library('dplyr')
library('forcats')
library('ggplot2')
library('ggrepel')
library('sjPlot')
library('plotROC')
library('rsample')
library('glmnet')
library('glmnetUtils')
library('vtable')
library('yardstick')
library('caret')
library('arm')
library('GGally')
library('repr')

```

```{r}

bowling_IPL <- read.csv("bowling_IPL.csv")
batting_IPL <- read.csv("batting_IPL.csv")

for (row in 1:nrow(bowling_IPL)) {
  num <- str_split_fixed(bowling_IPL$BBI[row], "/", n = 2)
  bbi = as.numeric(unlist(num[1])) / as.numeric(unlist(num[2]))
  bowling_IPL$BBI[row] <- round(bbi, digits = 2)
}

bowling_IPL <- bowling_IPL %>%
  mutate(bowling_IPL, 
          BBI = as.numeric(BBI),
          top_player = 0) %>%
  filter(Mat > 4)

bowling_IPL <- bowling_IPL %>%
  mutate(bowling_IPL,
         top_player = ifelse((
          (Runs < mean(bowling_IPL$Runs, na.rm = TRUE) + 2*sd(bowling_IPL$Runs)) & 
          (Wkts > mean(bowling_IPL$Wkts, na.rm = TRUE) - sd(bowling_IPL$Wkts)) & 
          (Avg < mean(bowling_IPL$Avg, na.rm = TRUE) + sd(bowling_IPL$Avg)) & 
          (Econ < mean(bowling_IPL$Econ, na.rm = TRUE) + sd(bowling_IPL$Econ)) & 
          (SR < mean(bowling_IPL$SR, na.rm = TRUE) + sd(bowling_IPL$SR)) &
          (BBI < mean(bowling_IPL$BBI, na.rm = TRUE) + sd(bowling_IPL$BBI))
          ), 1, 0)) %>%
  filter(Mat > 4)

bowling_IPL[is.na(bowling_IPL)] <- 0

bowling_IPL <- bowling_IPL %>%
  mutate(top_player = as.factor(top_player))

batting_IPL <- batting_IPL %>%
  mutate(batting_IPL, 
         HS = as.numeric(gsub("\\*", "", HS)),
         Avg = ifelse(Avg == '-', 0, as.numeric(Avg)),
         top_player = ifelse((
        (Runs < mean(batting_IPL$Runs, na.rm = TRUE)) & 
        (HS < mean(batting_IPL$HS, na.rm = TRUE)) &
        (Avg < mean(batting_IPL$Avg, na.rm = TRUE)) & 
        (SR < mean(batting_IPL$SR, na.rm = TRUE)) & 
        (X4s < mean(batting_IPL$X4s, na.rm = TRUE))
        ), 0, 1)) %>%
  filter(Mat > 4)

batting_IPL[is.na(batting_IPL)] <- 0

batting_IPL <- batting_IPL %>%
  mutate(top_player = as.factor(top_player))

bowling_IPL <- bowling_IPL[complete.cases(bowling_IPL),]
batting_IPL <- batting_IPL[complete.cases(batting_IPL),]

batting_split <- initial_split(batting_IPL, prop = 0.75)
batting_train <- training(batting_split)
batting_test <- testing(batting_split)

bowling_split <- initial_split(bowling_IPL, prop = 0.75)
bowling_train <- training(bowling_split)
bowling_test <- testing(bowling_split)

view(batting_IPL)
view(bowling_IPL)
```

Statistics for IPL Bowlers

```{r}

st(bowling_IPL)

```

Statistics for IPL Batters

```{r}

st(batting_IPL)

```

Plotting Top Player Comparison

```{r}

ggplot(data = bowling_IPL, aes(x = top_player, y = Wkts, group = top_player)) +
  geom_boxplot() + labs(title = 'Amount of Wickets Taken versus Top Player Status for Bowlers', x = 'Top Player', y = 'Amount of Wickets')

ggplot(data = batting_IPL, aes(x = top_player, y = Runs, group = top_player)) +
  geom_boxplot() + labs(title = 'Amount of Runs versus Top Player Status for Batters', x = 'Top Player', y = 'Amount of Runs')

```


Logistic Regression Model for Predicting Top Player Statistics

```{r}

bowlingLm <- glm(top_player ~ Mat + Ov + Runs + Wkts + Avg + Econ + BBI + SR,
                 family = binomial,
                data = bowling_train)

battingLm <- glm(top_player ~ Mat + NO + Avg + HS + Runs + BF + SR + X4s + X6s,
                family = binomial,
              data = batting_train)

battingBayes <- bayesglm(top_player ~ Mat + NO + Avg + HS + Runs + BF + SR + X4s + X6s,
                family = binomial(link = "logit"),
                data = batting_train)

```

Bowling LM Summary

```{r}

summary(bowlingLm)

exp(bowlingLm$coefficients)

```


Batting LM Summary

```{r}
options("scipen"=100)
summary(battingLm)

exp(battingLm$coefficients)

summary(battingBayes)

exp(battingBayes$coefficients)

```

Making Test and Train Prediction Tibbles

```{r}

bowling_preds_train <- predict(bowlingLm, newdata = bowling_train, type="response")

bowling_preds_test <- predict(bowlingLm, newdata = bowling_test, type="response")

bowling_results_train <- 
  tibble(
    `cmtruth` = as.factor(bowling_train$top_player),
    `roctruth` = as.numeric(bowling_train$top_player),
    `Class1` = bowling_preds_train,
    `Class2` = 1 - bowling_preds_train,
    `predicted` = as.factor(ifelse(bowling_preds_train > 0.4, 1, 0)))

bowling_results_test <- 
  tibble(
    `cmtruth` = as.factor(bowling_test$top_player),
    `roctruth` = as.numeric(bowling_test$top_player),
    `Class1` = bowling_preds_test,
    `Class2` = 1 - bowling_preds_test,
    `predicted` = as.factor(ifelse(bowling_preds_test > 0.4, 1, 0)))


batting_preds_train <- predict(battingLm, newdata = batting_train, type = "response")

batting_preds_test <- predict(battingLm, newdata = batting_test, type = "response")

bayes_preds_train <- predict(battingBayes, newdata = batting_train, type = "response")

bayes_preds_test <- predict(battingBayes, newdata = batting_test, type = "response")

batting_results_train <- 
  tibble(
    `cmtruth` = as.factor(batting_train$top_player),
    `roctruth` = as.numeric(batting_train$top_player),
    `Class1` = batting_preds_train,
    `Class2` = 1 - batting_preds_train,
    `predicted` = as.factor(ifelse(batting_preds_train > 0.4, 1, 0)))

batting_results_test <- 
  tibble(
    `cmtruth` = as.factor(batting_test$top_player),
    `roctruth` = as.numeric(batting_test$top_player),
    `Class1` = batting_preds_test,
    `Class2` = 1 - batting_preds_test,
    `predicted` = as.factor(ifelse(batting_preds_test > 0.4, 1, 0)))

bayes_results_train <- 
  tibble(
    `cmtruth` = as.factor(batting_train$top_player),
    `roctruth` = as.numeric(batting_train$top_player),
    `Class1` = batting_preds_train,
    `Class2` = 1 - batting_preds_train,
    `predicted` = as.factor(ifelse(bayes_preds_train > 0.4, 1, 0)))

bayes_results_test <- 
  tibble(
    `cmtruth` = as.factor(batting_test$top_player),
    `roctruth` = as.numeric(batting_test$top_player),
    `Class1` = batting_preds_test,
    `Class2` = 1 - batting_preds_test,
    `predicted` = as.factor(ifelse(bayes_preds_test > 0.4, 1, 0)))



```

Making ROC Graphs for Bowling and Batting Predictions

```{r}

bowling_train_roc <- ggplot(bowling_results_train, 
            aes(m = Class1, d = roctruth)) + 
  geom_roc(labelsize = 3.5, 
           cutoffs.at = 
             c(0.99,0.9,0.7,0.5,0.3,0.1,0)) +
  theme_minimal(base_size = 16)

bowling_test_roc <- ggplot(bowling_results_test, 
            aes(m = Class1, d = roctruth)) + 
  geom_roc(labelsize = 3.5, 
           cutoffs.at = 
             c(0.99,0.9,0.7,0.5,0.3,0.1,0)) +
  theme_minimal(base_size = 16)

batting_train_roc <- ggplot(batting_results_train, 
            aes(m = Class1, d = roctruth)) + 
  geom_roc(labelsize = 3.5, 
           cutoffs.at = 
             c(0.99,0.9,0.7,0.5,0.3,0.1,0)) +
  theme_minimal(base_size = 16)

batting_test_roc <- ggplot(batting_results_test, 
            aes(m = Class1, d = roctruth)) + 
  geom_roc(labelsize = 3.5, 
           cutoffs.at = 
             c(0.99,0.9,0.7,0.5,0.3,0.1,0)) +
  theme_minimal(base_size = 16)

bayes_train_roc <- ggplot(bayes_results_train, 
            aes(m = Class1, d = roctruth)) + 
  geom_roc(labelsize = 3.5, 
           cutoffs.at = 
             c(0.99,0.9,0.7,0.5,0.3,0.1,0)) +
  theme_minimal(base_size = 16)

bayes_test_roc <- ggplot(bayes_results_test, 
            aes(m = Class1, d = roctruth)) + 
  geom_roc(labelsize = 3.5, 
           cutoffs.at = 
             c(0.99,0.9,0.7,0.5,0.3,0.1,0)) +
  theme_minimal(base_size = 16)

```

Making Confusion Matrices for Bowling and Batting Predictions

```{r}

bowling_train_cm <- conf_mat(bowling_results_train,
                            truth = cmtruth,
                            estimate = predicted)

bowling_test_cm <- conf_mat(bowling_results_test,
                            truth = cmtruth,
                            estimate = predicted)

batting_train_cm <- conf_mat(batting_results_train,
                             truth = cmtruth,
                             estimate = predicted)
batting_test_cm <- conf_mat(batting_results_test,
                            truth = cmtruth,
                            estimate = predicted)

bayes_train_cm <- conf_mat(bayes_results_train,
                             truth = cmtruth,
                             estimate = predicted)
bayes_test_cm <- conf_mat(bayes_results_test,
                            truth = cmtruth,
                            estimate = predicted)

```


Bowling ROC Graphs

```{r}

print(bowling_train_roc)

print(bowling_test_roc)

```

Batting ROC Graphs

```{r}

print(batting_train_roc)

print(batting_test_roc)

print(bayes_train_roc)

print(bayes_test_roc)

```

Bowling AUC Scores

```{r}

calc_auc(bowling_train_roc)

calc_auc(bowling_test_roc)

```

Batting AUC Scores

```{r}

calc_auc(batting_train_roc)

calc_auc(batting_test_roc)

calc_auc(bayes_train_roc)

calc_auc(bayes_test_roc)

```

Bowling Prediction Confusion Matrices

```{r}

autoplot(bowling_train_cm, "heatmap")

autoplot(bowling_test_cm, "heatmap")

```

Batting Prediction Confusion Matrices

```{r}

autoplot(batting_train_cm, "heatmap")

autoplot(batting_test_cm, "heatmap")

autoplot(bayes_train_cm, "heatmap")

autoplot(bayes_test_cm, "heatmap")

```

Ridge Model Creation For Bowling

```{r}
ridge_bowling <- cv.glmnet(Econ ~ Mat + Ov + Runs + Wkts + Avg + Runs + BBI + SR,
                       data = bowling_train,
                       alpha = 0)

ridgeBow_mse <- ridge_bowling$cvm[ridge_bowling$lambda == ridge_bowling$lambda.min]
print(ridgeBow_mse)
print(ridge_bowling$lambda.min)

```

Plot

```{r}
plot(ridge_bowling)
```


Coefficients Of Lambdamin and Lambda.1se

```{r}
print(coef(ridge_bowling, s = ridge_bowling$lambda.min) %>% 
  round(3))

print(coef(ridge_bowling, s = ridge_bowling$lambda.1se) %>% 
  round(3))

```


For Batting

```{r}
ridge_batting <- cv.glmnet(Runs ~ Mat + NO + Avg + HS + Runs + BF + SR + X4s + X6s,
                       data = batting_train,
                       alpha = 0)

ridgeBat_mse <- ridge_batting$cvm[ridge_batting$lambda == ridge_batting$lambda.min]
print(ridgeBat_mse)
print(ridge_batting$lambda.min)

```

Plot

```{r}
plot(ridge_batting)
```
Coefficients Of Lambdamin and Lambda.1se

```{r}
print(coef(ridge_batting, s = ridge_batting$lambda.min) %>% 
  round(3))

print(coef(ridge_batting, s = ridge_batting$lambda.1se) %>% 
  round(3))


```


```{r}

plot(ridge_batting)

```


Lasso Model Creation For Bowling

```{r}

lasso_bowling <- cv.glmnet(Econ ~ Mat + Ov + Runs + Wkts + Avg + BBI + SR,
                       data = bowling_train,
                       alpha = 1)

lassoBow_mse <- lasso_bowling$cvm[lasso_bowling$lambda == lasso_bowling$lambda.min]
print(lassoBow_mse)
print(lasso_bowling$lambda.min)

```


Plot Lasso Model

```{r}
plot(lasso_bowling)
```


Coefficients Of Lambdamin and Lambda.1se

```{r}

print(coef(lasso_bowling, s = lasso_bowling$lambda.1se))
print(coef(lasso_bowling, s = lasso_bowling$lambda.min))

```


Lasso Model Creation For Batting

```{r}

lasso_batting <- cv.glmnet(Runs ~ Mat + NO + Avg + HS + Runs + BF + SR + X4s + X6s,
                       data = batting_train,
                       alpha = 1)

lassoBat_mse <- lasso_batting$cvm[lasso_batting$lambda == lasso_batting$lambda.min]
print(lassoBat_mse)
print(lasso_batting$lambda.min)

```


Plot Lasso Model

```{r}
plot(lasso_batting)
```

Coefficients Of Lambdamin and Lambda.1se

```{r}

print(coef(lasso_batting, s = lasso_batting$lambda.1se))
print(coef(lasso_batting, s = lasso_batting$lambda.min))

```

Minlossplots

```{r}
bowling_enet <- cva.glmnet(Econ ~ Mat + Ov + Runs + Wkts + Avg + Runs + BBI + SR,
                       data = bowling_train,
                       alpha = seq(0,1, by = 0.05))

batting_enet <- cva.glmnet(Runs ~ Mat + NO + Avg + HS + Runs + BF + SR + X4s + X6s,
                       data = batting_train,
                       alpha = seq(0,1, by = 0.05))

minlossplot(bowling_enet, cv.type="min")

minlossplot(batting_enet, cv.type="min")

```

Training vs Testing Set MSEs for Lasso Models

```{r}

lassoBowl_pred <- predict(lasso_bowling, s = lasso_bowling$lambda.min, newdata = bowling_test, type = "response")
Bowltest_mse <- mean((lassoBowl_pred - bowling_test$Econ)^2)

lassoBat_pred <- predict(lasso_batting, s = lasso_batting$lambda.min, newdata = batting_test, type = "response")
Battest_mse <- mean((lassoBat_pred - batting_test$Runs)^2)


cat("Lasso Bowling Model MSEs: \n")
cat("Training Set: ", lassoBow_mse, "\n")
cat("Testing Set: ", Bowltest_mse, "\n")
cat("\n")
cat("Lasso Batting Model MSEs: \n")
cat("Training Set: ", lassoBat_mse, "\n")
cat("Testing Set: ", Battest_mse, "\n")

```
